### 2013-05-24 Publications/2013-TMI/04-methods.tex
Added some primes in (12)
I have to include the Jacobian here.

\section{Methods}
\label{sec:methods}
%
\subsection{Maximum A-Posteriori Model}
\label{sec:methods_map}
%
A widely used approach to image segmentation is derived from the
Bayes' rule (\autoref{eq:bayes_rule}), where one seeks for a partitioning
of a certain image $X$ in piecewise smooth regions $\Omega = \lbrace \Omega_k , 
k\in\left[ 1 .. K \right] \rbrace$,  that maximizes the a posteriori 
probability given the multivariate image $X \in \mathbb{R}^C$, 
with $C$ being the number of image channels.
\begin{equation}
p(Y \mid X) = \frac{p(X \mid Y)\, p(Y)}{p(X)}.
\label{eq:bayes_rule}
\end{equation}

@@

A second widely-accepted assumption is the multivariate normal 
distribution of the different tissues in \gls{mri} data. Therefore,
the posterior probability of an infinitesimal voxel
can be written as:
\begin{equation}
p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) ) = \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{X(\mathbf{x})}) \right)}}.
\label{eq:bayes_mpdf}
\end{equation}
where we can identify the factor in the exponential, with 
$\mathbf{f} = X(\mathbf{x})$ as the squared \emph{Mahalanobis 
distance}:
\begin{equation}
\Delta^2_k (\mathbf{f}) = (\mathbf{f} - \boldsymbol{\mu}_k)^T \, \Sigma^{-1}_k \, (\mathbf{f} - \boldsymbol{\mu}_k).
\label{eq:bayes_mahalanobis}
\end{equation}

Finally, we can turn the \gls{map} problem into a variational one
applying the following log-transform:
\begin{align}
E(X \mid Y)= -\log \left[ p(X \mid Y) \, p(Y) \right] = \\
= -\log \left[ \underset{k}{\prod} \underset{\mathbf{x}\in \Omega_k}{\prod}
p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) )^{d\mathbf{x}} \right] = \\
= \sum\limits_k \int_{\Omega_k} -\log \left[ p_k(X(\mathbf{x}) \mid Y(\mathbf{x} ) ) \right] \, d\mathbf{x},
\label{eq:energy_1}
\end{align}
and introducing the posterior probability term (\autoref{eq:bayes_mpdf}), we get:
\begin{align}
E(X \mid Y) = \sum\limits_k \int_{\Omega_k} -\log \left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{f}) \right)}} \right] \, d\mathbf{x} = \\
= \sum\limits_k \int_{\Omega_k} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}.
\end{align}
Finally, after removing scaling factors and independent constants,
we obtain
\begin{align}
E(X \mid Y) = \sum\limits_k \left[ \log{\left|\boldsymbol{\Sigma}_{k}\right|} \int_{\Omega_k} d\mathbf{x} + \int_{\Omega_k} \Delta^2_k (\mathbf{f}) \,d\mathbf{x} \right],
\label{eq:map_energy}
\end{align}
(\textbf{FIXME:} bad explanation) where we have a constant term scaled by
the total volume of the partition $\Omega_k$ and the determinant of the 
covariance matrix of the partition $\left|\boldsymbol{\Sigma}_{k}\right|$,
plus an energy term based on the squared \emph{Mahalanobis distance}.

\subsection{Deformation model}
\label{sec:deformation}
%
The segmentation problem is transformed into a registration one if the initial
partition $Y$ is derived from the initial shape-priors in reference space. 
Introducing a dense deformation field $u$ that maps the original partition to
a better fit of the regions in the target coordinate space. Thus, the minimization
problem stated in \autoref{eq:map_energy} can be expressed so that $u$ is the 
unknown:
\begin{equation}
E(u(\mathbf{x})) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}
\label{eq:map_energy_deformation}
\end{equation}

>>>>

\section{Methods}
\label{sec:methods}
%
\subsection{Mumford-Shah Functional derivation from the Maximum A-Posteriori Model}
\label{sec:methods_map}
%
A widely used approach to image segmentation is derived from the
Bayes' rule \eqref{eq:bayes_rule}, where one seeks for a partitioning
of a certain image $X$ in piecewise smooth regions $\Omega = \lbrace \Omega_k , 
k\in\left[ 1 .. K \right] \rbrace$,  that maximizes the a posteriori 
probability given the multivariate image $X \in \mathbb{R}^C$, 
with $C$ being the number of image channels.
\begin{equation}
p(Y \mid X) = \frac{p(X \mid Y)\, p(Y)}{p(X)}.
\label{eq:bayes_rule}
\end{equation}

@@

A second widely-accepted assumption is the multivariate normal 
distribution of the different tissues in \gls{mri} data. Therefore,
the posterior probability of an infinitesimal voxel
can be written as:
\begin{equation}
p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) ) = \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{X(\mathbf{x})}) \right)}}.
\label{eq:bayes_mpdf}
\end{equation}
where we can identify the factor in the exponential, with 
$\mathbf{f} = X(\mathbf{x})$ as the squared \emph{Mahalanobis 
distance} with the parameters set $\Theta_k = \lbrace \boldsymbol{\mu}_k, \boldsymbol{\Sigma}_k \rbrace$:
\begin{equation}
\Delta^2_k (\mathbf{f} \mid \Theta_k ) = (\mathbf{f} - \boldsymbol{\mu}_k)^T \, \boldsymbol{\Sigma}^{-1}_k \, (\mathbf{f} - \boldsymbol{\mu}_k).
\label{eq:bayes_mahalanobis}
\end{equation}

Finally, we can turn the \gls{map} problem into a variational one
applying the following log-transform:
\begin{multline}
E(X \mid Y)= -\log \left[ p(X \mid Y) \, p(Y) \right] = \\
= -\log \left[ \underset{k}{\prod} \underset{\mathbf{x}\in \Omega_k}{\prod}
p_k( X(\mathbf{x}) \mid Y(\mathbf{x}) )^{d\mathbf{x}} \right] = \\
= \sum\limits_k \int_{\Omega_k} -\log \left[ p_k(X(\mathbf{x}) \mid Y(\mathbf{x} ) ) \right] \, d\mathbf{x},
\label{eq:energy_1}
\end{multline}
and introducing the posterior probability term \eqref{eq:bayes_mpdf}, we get:
\begin{multline}
E(X \mid Y) = \sum\limits_k \int_{\Omega_k} -\log \left[ \frac{1}{ \sqrt{(2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right|}}\,{e^{\left(-\frac{1}{2}  \Delta^2_k (\mathbf{f}) \right)}} \right] \, d\mathbf{x} = \\
= \sum\limits_k \int_{\Omega_k} \left[ \frac{1}{2} \log{ \left( (2\pi)^{C}\,\left|\boldsymbol{\Sigma}_{k}\right| \right)} + \frac{1}{2}  \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}.
\end{multline}
Finally, after removing scaling factors and independent constants,
we can express the functional in terms of $\lbrace\Theta_k,\Omega_k\rbrace$:
\begin{align}
E(\Theta_k,K) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x},
\label{eq:map_energy}
\end{align}
(\textbf{FIXME:} bad explanation) where we have a constant term scaled by
the total volume of the partition $\Omega_k$ and the determinant of the 
covariance matrix of the partition $\left|\boldsymbol{\Sigma}_{k}\right|$,
plus an energy term based on the squared \emph{Mahalanobis distance}.

Equation \eqref{eq:map_energy} resembles the Mumford-Shah functional including variance,
that modifies the original functional in a way that it can deal with more general
distributions. This is necessary to avoid the assumption that regions $\Omega_k$
have a fixed covariance matrix on their complete domain. One immediate advantage
of this functional from the original one is the possibility to distinguish regions
with the same mean vector but different covariance matrix \citep{brox_local_2009}:
\begin{multline}
E(\Theta_k,\Omega_k) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x} \\
+ \lambda \int_{\Omega_k - K}  ( \left| \nabla \mathbf{\mu} \right| ^2 + \left| \nabla \mathbf{\Sigma}_k \right| ^2 ) \, d\mathbf{x} 
+ \nu |K|,
\end{multline}
that is easily identifiable with \eqref{eq:map_energy} when we apply 
the so-called \emph{cartoon limit}, 
for $\lambda \to \infty$:
\begin{equation}
E(\Theta_k,K) = \sum\limits_k \int_{\Omega_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f}) \right] \,d\mathbf{x}
+ \nu |K|.
\end{equation}

\subsection{Deformation model}
\label{sec:deformation}
%
The segmentation problem is transformed into a registration one if the initial
partition $Y$ is derived from the initial shape-priors in reference space. 
Introducing a dense deformation field $u$ that maps the original partition to
a better fit of the regions in the target coordinate space. Thus, the minimization
problem stated in \eqref{eq:map_energy} can be expressed so that $u$ is the 
unknown:
\begin{equation}
E(u(\mathbf{x})) = \sum\limits_k \int_{\Omega'_k} \left[ \log \left|\mathbf{\Sigma}_k\right| + \Delta^2_k (\mathbf{f'}) \right] \,d\mathbf{x}
\label{eq:map_energy_deformation}
\end{equation}
where $\Omega'_k = u(\Omega_k)$ and $\mathbf{f'} = X(u(\mathbf{x}))$.
